
# Bivariate Analysis

**Figure 4 : Correlation Matrix on imputed Data**
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.width = 13, fig.height = 13}
get_correlation_matrix <- function(data, Outcome = NULL, use = "pairwise.complete.obs") {

  # Select columns based on the presence of ID and Outcome
  if ("ID" %in% colnames(data)) {
    if (!is.null(Outcome)) {
      cor_matrix <- cor(data %>% select(-ID, -Outcome), use = use)
    } else {
      cor_matrix <- cor(data %>% select(-ID), use = use)
    }
  } else {
    if (!is.null(Outcome)) {
      cor_matrix <- cor(data %>% select(-Outcome), use = use)
    } else {
      cor_matrix <- cor(data, use = use)
    }
  }

  # Create the heatmap to visualize the correlation matrix
  ggplot(data = as.data.frame(as.table(cor_matrix)), aes(Var1, Var2, fill = Freq)) +
    geom_tile() +
    geom_text(aes(label = round(Freq, 2)), color = "black", size = 3) +
    scale_fill_gradient2(low = "#3838ff", high = "#ff4444", mid = "#ffffff", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
    coord_fixed() +
    labs(title = "", x = "", y = "")
}

get_correlation_matrix(data)

```

On this figure, the values of the Pearson 
correlation coefficients (noted α) are indicated for each pair of variables.  
Examination of these coefficients shows:  
1. '`STDsHPV`' ≡ 'STDsVaginalCondylomatosis' , 'STDsPelvicInflammatoryDisease' , 'STDsGenitalHerpes' , 'STDsMolluscumContagiosum' , 'STDsHIV' , 'STDsHepatitisB' (α < 0.90)  
2. '`STDsHepatitisB`' ≡ 'STDsVaginalCondylomatosis' , 'STDsPelvicInflammatoryDisease' , 'STDsGenitalHerpes' , 'STDsMolluscumContagiosum' , 'STDsHIV' , 'STDsHPV' (α < 0.90)  
3. '`STDsHIV`' ≡ 'STDsPelvicInflammatoryDisease' , 'STDsGenitalHerpes' , 'STDsMolluscumContagiosum' , 'STDsHepatitisB' , 'STDsHPV' (α < 0.90)  
4. '`STDsMolluscumContagiosum`' ≡ 'STDsVaginalCondylomatosis' , 'STDsPelvicInflammatoryDisease' , 'STDsGenitalHerpes' , 'STDsHPV' , 'STDsHIV' , 'STDsHepatitisB' (α < 0.90)  
5. '`STDsGenitalHerpes`' ≡ 'STDsVaginalCondylomatosis' , 'STDsPelvicInflammatoryDisease' , 'STDsMolluscumContagiosum' , 'STDsHPV' , 'STDsHIV' , 'STDsHepatitisB' (α < 0.90)  
6. '`STDsPelvicInflammatoryDisease`' ≡ 'STDsVaginalCondylomatosis' , 'STDsGenitalHerpes' , 'STDsMolluscumContagiosum' , 'STDsHPV' , 'STDsHIV' , 'STDsHepatitisB' (α < 0.90)  
7. '`STDsVaginalCondylomatosis`' ≡ 'STDsPelvicInflammatoryDisease' , 'STDsGenitalHerpes' , 'STDsMolluscumContagiosum' , 'STDsHPV' , 'STDsHIV' , 'STDsHepatitisB' (α < 0.90)  
8. '`STDsNumber`' ≡ 'STDs' (α < 0.90) 

There are many variables that seem to have a linear relationship in the data frame

## Quantitative VS Outcome

Outcome = `DxCancer`

### Vizualisation
#### Boxplots by diagnosis
```{r Quantitative VS Outcome plots ,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4}

quantitative <- colnames(data[colnames(data) %in% quantitative_variable_names])

# Boucle pour générer des scatterplots indépendants pour chaque paire de variables
plot_list <- list()  # Créer une liste pour stocker les graphiques

for (i in 1:(length(quantitative)-1)) {
  
  var_x <- Outcome  # La variable Outcome
  var_y <- quantitative[i]  # Prendre chaque variable quantitative

  if (var_x != var_y) {
    # Créer le boxplot avec coloration en fonction de Outcome
    p <- ggplot(data, aes(x = .data[[var_x]], y = .data[[var_y]], fill = .data[[var_x]])) +
      geom_boxplot() +
      scale_fill_manual(values = c("TRUE" = "#83b9cb", "FALSE" = "#f49691")) +  # Colorer TRUE en bleu, FALSE en rouge
      labs(title = paste("Boxplot of", var_y, "by", var_x)) +
      theme_minimal() +
      theme(plot.title = element_text(size = 10),
            plot.margin = unit(c(1, 1, 3, 1), "lines"),
            legend.position = "none")  # Supprimer la légende si non nécessaire

    # Ajouter le plot à la liste
    plot_list[[paste(var_y, var_x, sep = "_by_")]] <- p
  }
}

# Afficher les graphiques (par exemple, 2 colonnes de graphiques)
aes_columns_display(plot_list, 2)
```



#### Density plots by diagnosis

```{r Quantitative VS Outcome density ,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4}

density_plot_list <- list()  # Créer une liste pour stocker les graphiques

for (i in 1:length(quantitative)) {
  
    var_x <- quantitative[[i]]  # Extraire la ième variable quantitative
    var_y <- Outcome  

    if (!identical(var_x, var_y)) {  # Assurez-vous que la variable quantitative n'est pas Outcome
      
        # Créer le graphique
        p <- ggplot(data, aes_string(x = var_x, fill = var_y, colour = var_y)) +
                  geom_density(alpha = 0.25)+
          geom_rug() +
          labs(title = paste("Density plot of", var_x, "by Cancer diag"),
               x = names(quantitative)[i],
               fill = "Cancer diag",
               colour = "Cancer diag")+
      theme_minimal() + 
      theme(plot.title = element_text(size = 10))
        
   density_plot_list[[paste(var_y, var_x, sep = "_by_")]] <- p
    }
    else {
       
    }
  }


# Afficher les graphiques (par exemple, 2 colonnes de graphiques)
aes_columns_display(density_plot_list, 2)
```

### Statistics

This statistical section will allow us to observe whether there is a 
significant difference between patients diagnosed with cervical cancer and 
undiagnosed patients according to the different quantitative variables in the study. 
We begin by testing the normality of the variables to find out the appropriate statistical test to use. 

#### Normality test for continuous variables
```{r Shapiro test ,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4}

shapiro_list <- list()

for (i in 1:length(continuous_variable_names)) {
var <- continuous_variable_names[[i]]
res <- shapiro.test(data[[var]])

res <- shapiro.test(data[[var]])
  cat("\nShapiro Test Result for", var, "distribution :\n")
  print(res)
  cat("\n")

}

```

#### QQplot
```{r qq visualisation ,echo = FALSE, message=FALSE, warning=FALSE, results='hide', fig.height = 4, fig.width=5}

visual_list <- list()

for (i in 1:length(continuous_variable_names)) {
var <- continuous_variable_names[[i]]

p<-qqnorm(data[[var]], main = NULL)
  qqline(data[[var]])
  title(main = paste(var))
}

```

None of the distributions is normal. The Wilcoxon test will be used.

#### tests

The wilcoxon test will allow us to establish whether there is a significant difference between 
patients diagnosed with cancer and those not diagnosed for the different quantitative variables in the study.

```{r wilcoxon tests ,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 10}

wilcox_list <- list()

for (i in 1:length(continuous_variable_names)) {
var <- continuous_variable_names[[i]]

res <- wilcox.test(data[[var]]~data[[Outcome]])
wilcox_list[[paste(var)]] <- res
}

wilcox_list
```

The statistical test assumes that there is a statistically significant difference between the groups for age, IUDYears (Years of IUD use) (p-value < 0.05)  


## Qualitative VS Outcome

### Vizualisation and statistics

The Pearson's Chi-squared test will allow us to establish whether there is a significant difference between 
patients diagnosed with cancer and those not diagnosed for the different qualitative variables in the study.

```{r Qualitative VS Outcome plots,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4}

qualitative <- colnames(data[colnames(data) %in% qualitative_variable_names])

# Créer une liste pour stocker les graphiques
plot_list <- list()

# Boucle pour générer des barplots indépendants pour chaque paire de variables
for (i in 1:length(qualitative)) {
  
    var_x <- qualitative[i]  # Variable qualitative
    var_y <- Outcome         # Outcome (TRUE/FALSE)

    if (var_x != var_y){
      # Créer un barplot avec les proportions de Outcome par catégorie de var_x
      p <- ggplot(data, aes(x = .data[[var_x]], fill = factor(.data[[var_y]]))) +
        geom_bar(position = "fill") +
        scale_y_continuous(labels = scales::percent_format()) +  # Pour afficher en pourcentage
        labs(title = paste("Proportion of", var_y, "by", var_x), 
             x = var_x, y = "Proportion", fill = var_y) +
        theme_minimal() +
        theme(plot.title = element_text(size = 10),
            plot.margin = unit(c(1, 1, 3, 1), "lines"))
      
      # Ajouter le plot à la liste
      plot_list[[paste(var_x, var_y, sep = "_by_")]] <- p
    }
}

```


```{r Chi2 tests ,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4}

chi2_list <- list()

for (i in 1:length(qualitative)) {
  
    var_x <- qualitative[i]  # Variable qualitative
    var_y <- Outcome

table_res <- table(data[[var_x]], data[[var_y]])
res <- chisq.test(table_res)

 chi2_list[[paste(var_x, var_y, sep = "_by_")]] <- res
}

```


```{r Qualitative display ,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4, fig.width=5}

for (i in 1:length(qualitative)) {

  var_x <- qualitative[i]  # Variable qualitative
  var_y <- Outcome

  if (var_x != var_y){
  # Afficher le graphique
  p <- plot_list[[paste(var_x, var_y, sep = "_by_")]]
  print(p)
  
  # Afficher les résultats du test
  res <- chi2_list[[paste(var_x, var_y, sep = "_by_")]]
  cat("\nChi-Square Test Result for", var_x, "and", var_y, "association :\n")
  print(res)
  cat("\n")
}
}
```

The statistical test assumes that there is a statistically significant difference between the groups for IUD (IUD use), DxHPV, Dx, Hinselmann, Schiller, Cytology, Biopsy (p-value < 0.05)  
