# Multivariate Analysis

To improve the results of the multivariate analysis, data has been normalised.  

>`scale()` is used from the heatmaply library. description: `scale()` use Z-score.  
This method is robust to extreme values and retains information on the variance. 


## Vizualisation

```{r headmap on data, echo = FALSE, message=FALSE, warning=FALSE, fig.height = 15}
#headmap<- heatmaply(
#  data, 
#  xlab = "Features",
#  ylab = "Patients", 
#  main = "data headmap"
#)
```


```{r normalisation, echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4} 
normalised_data <- scale(data %>% select(-ID))
```

**Figure 5 : heatmap on standardised data**

```{r headmap on normalised data, echo = FALSE, message=FALSE, warning=FALSE, fig.height = 15}
#headmap_normalized<- heatmaply(
#  normalised_data, 
#  xlab = "Features",
#  ylab = "Patients", 
#  main = "Normalised data headmap"
#)
#headmap_normalized
```


**Figure 6 : Correlation matrix on standardised data**

```{r correlation matrix on normalised data, echo = FALSE, message=FALSE, warning=FALSE, fig.width = 13, fig.height = 13}

get_correlation_matrix(normalised_data)

```



## Linear Regression





## Logistic Regression





## Interpretation 

### OR 
### RR 