# Multivariate Analysis 

To improve the results of the multivariate analysis, i normalise the data. 
scale() is used from the heatmaply library. description: scale() normalise par rapport à la moyenne et à l'écart-type, ce qui donne un Z-score.  
I chose this method because: there are a few extreme values in my unsuppressed dataset (professional expertise) and this method is robust to 
outliers and retains information on the variance. 

## Vizualisation

```{r headmap on data, echo = FALSE, message=FALSE, warning=FALSE, fig.height = 15}
#headmap<- heatmaply(
#  data, 
#  xlab = "Features",
#  ylab = "Patients", 
#  main = "data headmap"
#)
```


```{r normalisation, echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4} 
normalised_data <- scale(data %>% select(-ID))
```

**Figure 5 : heatmap on standardised data**

```{r headmap on normalised data, echo = FALSE, message=FALSE, warning=FALSE, fig.height = 15}
#headmap_normalized<- heatmaply(
#  normalised_data, 
#  xlab = "Features",
#  ylab = "Patients", 
#  main = "Normalised data headmap"
#)
#headmap_normalized
```


**Figure 6 : Correlation matrix on standardised data**

```{r correlation matrix on normalised data, echo = FALSE, message=FALSE, warning=FALSE, fig.width = 13, fig.height = 13}

get_correlation_matrix(normalised_data)

```

## ACP

```{r ACP , echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4}

selected_columns <- c(which(colnames(normalised_data) %in% quantitative_variable_names), which(colnames(normalised_data) == "DxCancer"))

# Effectuer la PCA avec la variable "DxCancer" incluse comme qualitative supplémentaire
res_PCA <- PCA(normalised_data[, selected_columns], quali.sup = which(colnames(normalised_data[, selected_columns]) == "DxCancer"))

#explor::explor(res_PCA)
res_prepared <- explor::prepare_results(res_PCA)


res_var12 <- explor::PCA_var_plot(res_prepared, xax = 1, yax = 2, var_sup = FALSE, 
                                  var_sup_choice = c(), var_lab_min_contrib = 0, 
                                  col_var = NULL, labels_size = 10, scale_unit = TRUE, 
                                  transitions = TRUE, labels_positions = NULL)

res_var13 <- explor::PCA_var_plot(res_prepared, xax = 1, yax = 3, var_sup = FALSE, 
                                  var_sup_choice = c(), var_lab_min_contrib = 0, 
                                  col_var = NULL, labels_size = 10, scale_unit = TRUE, 
                                  transitions = TRUE, labels_positions = NULL, 
                                  xlim = c(-1.1, 1.1), ylim = c(-1.1, 1.1))

res_var23 <- explor::PCA_var_plot(res_prepared, xax = 2, yax = 3, var_sup = FALSE, 
                                  var_sup_choice = c(), var_lab_min_contrib = 0, 
                                  col_var = NULL, labels_size = 10, scale_unit = TRUE, 
                                  transitions = TRUE, labels_positions = NULL, 
                                  xlim = c(-1.1, 1.1), ylim = c(-1.1, 1.1))


res_ind <- explor::PCA_ind_plot(res_prepared, xax = 1, yax = 2, ind_sup = FALSE, 
                                lab_var = NULL, ind_lab_min_contrib = 0, col_var = "DxCancer", 
                                labels_size = 9, point_opacity = 0.5, opacity_var = NULL, 
                                point_size = 64, ellipses = FALSE, transitions = TRUE, 
                                labels_positions = NULL)


res_var12
res_var13
res_var23
res_ind
```

## ACM

```{r ACM, echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4}
selected_columns <- c(which(colnames(data) %in% qualitative_variable_names))

res_MCA <- MCA(data[, selected_columns])

#explor::explor(res_MCA)
res_prepared <- explor::prepare_results(res_MCA)

res_var12 <- explor::MCA_var_plot(res_prepared, xax = 1, yax = 2, var_sup = FALSE, var_sup_choice = ,
    var_lab_min_contrib = 0, col_var = "Variable", symbol_var = NULL, size_var = NULL,
    size_range = c(10, 300), labels_size = 10, point_size = 56, transitions = TRUE,
    labels_positions = NULL, labels_prepend_var = FALSE, xlim = c(-2.54, 4.09),
    ylim = c(-1.13, 5.5))

res_var13 <- explor::MCA_var_plot(res_prepared, xax = 1, yax = 3, var_sup = FALSE, var_sup_choice = ,
    var_lab_min_contrib = 0, col_var = "Variable", symbol_var = NULL, size_var = NULL,
    size_range = c(10, 300), labels_size = 10, point_size = 56, transitions = TRUE,
    labels_positions = NULL, labels_prepend_var = FALSE, xlim = c(-5.01, 7.18),
    ylim = c(-5.7, 6.49))

res_ind <- explor::MCA_ind_plot(res_prepared, xax = 1, yax = 2, ind_sup = FALSE, lab_var = NULL,
    ind_lab_min_contrib = 0, col_var = "DxCancer", labels_size = 9, point_opacity = 0.5,
    opacity_var = NULL, point_size = 64, ellipses = FALSE, transitions = TRUE,
    labels_positions = NULL)

res_var12
res_var13
res_ind

```

